A linearization procedure for constrained multibody systems

### 2012-04-21 PetersonGedeHubbard2012.tex
Finished putting in rolling disc equations.

\section{Example}
\label{example}
We apply this linearization technique to the well studied rolling disc.  Figure
XX illustrates the choice of coordinates, speeds, and parameters.  The disc is
assumed to be knife edged of radius $r$, mass $m$, spin inertia $J$, and
transverse inertia $I$.  To illustrate the concept of dependent quantities, we
purposefully choose a nonminimal set of coordinates and speeds.  In this
particular example, this can be avoided, but in more complicated systems, it is
difficult (if not impossible) to describe the system with a minimal choice.  The
orientation of the disc fixed frame $C$ relative to the inertial frame $N$ is given by
a sequence of angular rotations (Euler ZXY) by angles $q_0$ (yaw), $q_1$
(lean), $q_2$ (spin).  The location
of the center of disc relative to the inertial origin is defined by coordinates
q3, q4, q5.  The angular velocity of the disc is defined by three speeds
\[u_0 \triangleq {^{N}}\bm{\omega}^{C} \cdot \bm{c}_{x}\]

\section{Derivation of linearization algorithm}
\label{sec:derivations}

>>>>

\section{Example}
\label{example}
We apply this linearization technique to the well studied rolling disc.  Figure
XX illustrates the choice of coordinates, speeds, and parameters.  The disc is
assumed to be knife edged of radius $r$, mass $m$, spin inertia $J$, and
transverse inertia $I$.  To illustrate the concept of dependent quantities, we
purposefully choose to describe the configuration and velocity of the disc with
a non-minimal set of coordinates and speeds.  This choice is not purely
academic; a physical disc equipped with a rate gyroscope and accelerometer
measure these body fixed quantities (or their time derivatives) directly,
rather than the more traditional set which is chosen for analytic convenience.
In more complicated systems with constraints, it is difficult (if not
impossible) to describe the system with a minimal choice of coordinates and
speeds; the ideas demonstrated by this simply example apply analogously to more
complicated ones.

The orientation of the disc fixed frame $C$ relative to the inertial frame $N$
is given by a sequence of angular rotations (Euler ZXY) by angles $q_0$ (yaw),
$q_1$ (lean), $q_2$ (spin).  The angular velocity of $C$ in $N$, the location
of disc center $C^*$ relative to the inertial origin $N^*$, and the velocity of
$C^*$ in $N$ are
\begin{align}
  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
\end{align}
If the disc is to make contact with the ground, it should be obvious to the
reader that there is a configuration constraint involving the height $q_5$ of
the disc center and the lean $q_1$.  Similarly, if the disc rolls without
slip, it should be obvious that there are constraints which relate the
generalized speeds $u_i, (i = 0,\dots,5)$.  The derivation of these
constraints and the equations of motion are presented in the electronic
supplementary material and can be found in the literature (reference O'Reilly
and others?).
\begin{table}[htbp]
  \centering
  \caption{Rolling disc governing definitions and equations}
  \label{table:rollingdisc}
  \begin{tabular}[c]{l l l}
    Quantity & Shape & Description\\
    \hline
    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
    derivatives\\
    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
  \end{tabular}
  \begin{tabular}[c]{r @{ $=$ } l l}
    \multicolumn{3}{c}{ } \\
    \multicolumn{2}{c}{Equation} &           \\
    \hline
    $rc_1 + q_5$ & $0$ & $\bm{f}_c = \bm{0}$ \\
    $\left[\begin{array}{c}
        r u_1 c_2 + u_3 \\
        -r (u_0 c_2 + u_2 s_2) + u_4 \\
        r u_1 s_2 + u_5
    \end{array}\right] $ &
    $\left[\begin{array}{c}$0$\\$0$\\$0$\end{array}\right]$
      & $\bm{f}_v = \bm{0}$\\
    $\left[\begin{array}{c}
        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
      \end{array}\right]$ &
      $\left[\begin{array}{c}$0$\\$0$\\$0$\end{array}\right]$ & $\bm{f}_a =
        \bm{0}$\\
        $\left[\begin{array}{c}
-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
\dot{q}_0 s_1 + \dot{q}_2 \\
\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
- \dot{q}_5 s_2 c_1\\
- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
         \end{array}\right] + \left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ &
             $\left[\begin{array}{c} 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{array}\right]$
                 & $\bm{f}_0 + \bm{f}_1 = \bm{0}$\\
                 $\left[\begin{array}{c}
-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
-I \dot{u}_2 - m r s_2 \dot{u}_4
                   \end{array}\right]
+
                 \left[\begin{array}{c}
-I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
                m r (u_0 u_4 - u_1 u_3) s_2 + m r (u_1 u_5 - u_2 u_4) c_2 \\
 I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
                   \end{array}\right]$ &
             $\left[\begin{array}{c} 0 \\ 0 \\ 0 \end{array}\right]$
               & $\bm{f}_2 + \bm{f}_3 = \bm{0}$
  \end{tabular}
\end{table}

%f_c:
%[r*cos(q1) + q5]
%f_v:
%[                r*u1*cos(q2) + u3]
%[-r*(u0*cos(q2) + u2*sin(q2)) + u4]
%[                r*u1*sin(q2) + u5]
%f_a:
%[                                -r*u1*sin(q2)*q2' + r*cos(q2)*u1' + u3']
%[-r*(-u0*sin(q2)*q2' + u2*cos(q2)*q2' + sin(q2)*u2' + cos(q2)*u0') + u4']
%[                                 r*u1*cos(q2)*q2' + r*sin(q2)*u1' + u5']
%f_0:
%[                                                                                      -sin(q2)*cos(q1)*q0' + cos(q2)*q1']
%[                                                                                                       sin(q1)*q0' + q2']
%[                                                                                       sin(q2)*q1' + cos(q1)*cos(q2)*q0']
%[(sin(q0)*cos(q2) + sin(q1)*sin(q2)*cos(q0))*q4' + (-sin(q0)*sin(q1)*sin(q2) + cos(q0)*cos(q2))*q3' - sin(q2)*cos(q1)*q5']
%[                                                                -sin(q0)*cos(q1)*q3' + sin(q1)*q5' + cos(q0)*cos(q1)*q4']
%[ (sin(q0)*sin(q2) - sin(q1)*cos(q0)*cos(q2))*q4' + (sin(q0)*sin(q1)*cos(q2) + sin(q2)*cos(q0))*q3' + cos(q1)*cos(q2)*q5']
%f_1:
%[-u0]
%[-u1]
%[-u2]
%[-u3]
%[-u4]
%[-u5]
%f_2:
%[                  -I*u0' - m*r*cos(q2)*u4']
%[-J*u1' + m*r*sin(q2)*u5' + m*r*cos(q2)*u3']
%[                  -I*u2' - m*r*sin(q2)*u4']
%f_3:
%[-I*u1*u2 + J*u1*u2 + g*m*r*sin(q1)*cos(q2) - m*r*(-u0*u5 + u2*u3)*cos(q2)]
%[                m*r*(u0*u4 - u1*u3)*sin(q2) + m*r*(u1*u5 - u2*u4)*cos(q2)]
%[ I*u0*u1 - J*u0*u1 + g*m*r*sin(q1)*sin(q2) - m*r*(-u0*u5 + u2*u3)*sin(q2)]
\section{Derivation of linearization algorithm}
\label{sec:derivations}
